---
title: "Data Tidying"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(dplyr)
library(tidyr)
library(readr)

```

# Read in data
Data come from [Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2)


```{r}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", show_col_types = FALSE)

head(catch_original)

```

# Clean and reshape data

Drop notes and all column
```{r}
catch_data <- catch_original %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)

head(catch_data)
```

Coerce Chinooch column to be numeric instead of character

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook))

summary(catch_clean)
```


## Examine NA values
```{r}
i <-which(is.na(catch_clean$Chinook))
i
catch_original[i, ]
```

```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "species", values_to = "catch")

head(catch_long)

```

```{r}
var <- catch_long %>% 
  filter(Year > 1899) %>% 
  group_by(species, Year) %>% 
  summarise(mean_catch = mean(catch), .groups = "drop")

head(var)

```

{r}
mean_region <- catch_long %>% 
  group_by(species) %>% 
  summarise(mean_catch = mean(catch))


mean_region


```
```{r} 
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea)

head(region_defs)
```

